---
title: "`biocmask` Usage Guide"
author:
    - name: Justin Landis
      orcid: 0000-0001-5501-4934
    - name: Michael Love
      orcid: 0000-0001-8401-0545
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: rmarkdown::html_document
abstract: >
  biocmask is great.
bibliography: library.bib
vignette: |
  %\VignetteIndexEntry{biocmask Usage Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Quick start

One or two sentence overview of what is going on here...

## Supported data types and operations

`biocmask` works on *SummarizedExperiment* objects, as well as most
classes derived from this, including *DESeqDataSet*,
*SingleCellExperiment*, etc.

It supports the following operations:

* `mutate`
* `select`
* ...

## Typical use case

```{r message=FALSE}
library(airway)
data(airway)
library(dplyr)
library(biocmask)
airway |>
  mutate(log_counts = log1p(counts),
         cols(treated = dex == "trt"),
         rows(new_id = paste0("gene-", gene_name)))
```

## Related work

* these
* are 
* some 
* others

# Manipulating *SummarizedExperiment* with `biocmask`

## assay context

## row context

## col context

# How it works

# Advanced features

# Troubleshooting and best practices

# Community and support

# Session info

```{r}
devtools::session_info()
```

# References
